---
title: "Simulations OVB/type M trade-off: RDD"
author:
  - name: Vincent Bagilet 
    url: https://www.sipa.columbia.edu/experience-sipa/sipa-profiles/vincent-bagilet
    affiliation: Columbia University
    affiliation_url: https://www.columbia.edu/
  - name: Léo Zabrocki 
    url: https://www.parisschoolofeconomics.eu/en/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/en/
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE, results='hide', warning=FALSE}
library(knitr)
opts_chunk$set(fig.path = "images/",
               cache.path = "cache/",
               cache = FALSE,
               echo = TRUE, #set to false to hide code
               message = FALSE,
               warning = FALSE,
               out.width = "85%",
               dpi = 200,
               fig.align = "center")  
```  

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse) 
library(knitr) 
library(mediocrethemes)
library(fixest)
library(tictoc)

set_mediocre_all()
```

## Purpose of the document

In this document, we run a simulation exercise to illustrate the existence of a trade-off between Omitted Variable Bias (OVB) and type M error in the case of the Regression Discontinuity Design (RDD). This trade-off is mediated by the size of the bandwidth considered in the analysis. The idea behind this is that the smaller the bandwidth, the more comparable units are and therefore the smaller the risk of OVB is. Yet, with a smaller bandwidth, sample size, and thus power, decreases, increasing the risk of type M error. 

## Extremely simple example

We first implement a very simple example to get our hands on simulating a RDD with an omitted variable bias.

We set baseline values for all the parameters and generate the data. Note that the bandwidth is defined as a proportion of the standard deviation of the variable x.

```{r DGP}
generate_data_rdd <- function(n_indiv = 10000, 
                              n_period = 1,  
                              sigma_e = 0.4, 
                              mu_u = 1, 
                              sigma_u = 0.2, 
                              mu_x = 2, 
                              sigma_x = 4, 
                              intensity_ovb = 10, 
                              # delta_y = 10, 
                              alpha = 0, 
                              gamma = 0.3, 
                              beta = 1, 
                              sigma_t = 2,
                              bw = 0.1
                              ) {
  
  threshold_pos <- runif(1, 0.2, 0.8) 
  delta_x <- intensity_ovb
  delta_y <- intensity_ovb
  
  data <- 
    tibble(indiv = 1:n_indiv) %>% 
    crossing(period = 1:n_period) %>% 
    mutate(
      uniq_id = paste(indiv, period, sep = "_"),
      e = rnorm(n_indiv*n_period, 0, sigma_e),
      u = rnorm(n_indiv*n_period, mu_u, sigma_u),
      x = rnorm(n_indiv*n_period, mu_x, sigma_x)+ delta_x*u,
      Y0 = alpha + gamma*x + delta_y*u + e, 
      treated = .data[["x"]] > quantile(.data[["x"]], threshold_pos, names = FALSE),
      Y1 = Y0 + beta*treated + rnorm(nrow(.), 0, sigma_t), 
      in_bw = dplyr::between(x, mu_x - bw/2*sigma_x, mu_x + bw/2*sigma_x)
    ) 
  
  return(data)
}
```

We want to vary two parameters only: the bandwidth and the intensity of the omitted variable bias. The function `estimate_rdd` therefore only takes two parameters. 

```{r estimate}
# data_rdd_treated %>% 
#   ggplot(aes(x = x, y = u)) +
#   geom_point()

estimate_rdd <- function(intensity_ovb, bw) {
  lm(
    data = generate_data_rdd(intensity_ovb = intensity_ovb, bw = bw), 
    formula = Y1 ~ treated + x
  ) %>% 
  broom::tidy() %>% 
  filter(term == "treatedTRUE") %>%
  rename(p_value = p.value, se = std.error) %>%
  select(estimate, p_value, se) %>% 
  mutate(
    intensity_ovb = intensity_ovb,
    bw = bw
  )
}

# estimate_rdd(intensity_ovb = 5, bw = 0.1)
```

We then create a table with all the values of the parameters we want to test. We consider 

```{r simulate, eval=FALSE}
vect_bw <- c(0.05, 0.1, 0.2, 0.5, 1)
vect_intensity_ovb <- c(1, 10)
n_iter <- 100

param_rdd <- crossing(vect_bw, vect_intensity_ovb) %>% 
  rename(bw = vect_bw, intensity_ovb = vect_intensity_ovb) %>% 
  crossing(rep_id = 1:n_iter) %>% 
  select(-rep_id)

tic()
simulations_rdd <- pmap_dfr(param_rdd, estimate_rdd)
toc()
```











